---
title: "02 Phylogenetic Signal"
author: "Katie Heineman"
date: "6/7/2019"
output: html_document
---

# Analyzing Phylogenetic Signal in Germination Percent and Seed Time to Median Germination

This document also include visualization of the traits on the tree

```{r setup, include=FALSE}
# Packages
library(data.table)
library(picante)
library(MCMCglmm)

# Actual current version of ggtree not supported in this version of R, but this should work for installation
#source("https://bioconductor.org/biocLite.R")
#biocLite("ggtree")
library(ggtree)

# Ana's seed database
seedDat <- fread("Data/Savanna seed database.csv")
setnames(seedDat,c("Vegetation type", "Dispersal mode","Moisture content", "Seed mass (mg)"), c("VegType", "DM","MC","SM" ))
seedDat[,Dormancy2Cat:=ifelse(Dormancy=="ND","ND","D")]

## Read in Trees 
# Scenario 3:  adding genera or species to their families or genera using the same approach implemented in Phylomatic and BLADJ 
phy_coll3 <- read.tree("Data/phy_coll3.tre")
# Scenario 1:  adding genera or species as basal polytomies within their families or genera 
phy_coll1 <- read.tree("Data/phy_coll1.tre")
# Scenario 2: adding genera or species randomly within their families or genera 
phy_coll2 <- read.tree("Data/phy_coll2.tre")

```

## Data Prep for Phylogenetic Analysis
```{r data prep phylo}
# Prep seed dat by making species with underscore
seedDat[,Species:=gsub(" ","_",Species)]

# Calculating Species Means
seedDat_spp <- seedDat[,.(T50_mean=mean(T50,na.rm=T),log_T50_mean=mean(log(T50),na.rm=T),GP = mean(GP,na.rm=T), sqrt_GP = mean(sqrt(GP),na.rm=T),Seedmass = mean(SM,na.rm=T), Moisture = mean(MC,na.rm=T)), by="Species"]
# The number of sucess is the averager germ percentage X 10 seeds per replicate X 10 reps
seedDat_spp[,Successes:=GP]
seedDat_spp[,Failures:=100-Successes]

seedDat_spp1 <- copy(seedDat_spp)
# Make this dataset into a matrix
rowlab <- seedDat_spp1$Species
trait_matrix = as.matrix(seedDat_spp1[,Species:=NULL])
rownames(trait_matrix) = rowlab

```




## Calculating Phylogenetic Signal Using Picante Package

The statistic calculated here is the K statistic with a one tailed test for significance relative to random type reshuffling (https://cran.r-project.org/web/packages/picante/vignettes/picante-intro.pdf)

```{r old school phylogenetic signal, message=FALSE, warning=FALSE}

# Ensuring trait matrix order matches the phylogenetic tree order 
trait_matrix <- trait_matrix[phy_coll3$tip.label,]

# Creating a data object that matches the tree and the traits (only including T50 and GP)
MatchDataObj= match.phylo.data(phy_coll3, trait_matrix[,c(1,3)])

# Calculating phylogenetic signal as well as P-value based on variance of phylogenetically independent contrasts relative to tip shuffling randomization.
physig = multiPhylosignal(MatchDataObj$data, MatchDataObj$phy,reps=100000)

physig


# Testing if the "scenario matters" (Scenario 2 is different than 1 or 3)
# Ensuring trait matrix order matches the phylogenetic tree order 
trait_matrix <- trait_matrix[phy_coll2$tip.label,]

# Creating a data object that matches the tree and the traits (only including T50 and GP)
MatchDataObj= match.phylo.data(phy_coll2, trait_matrix[,c(1,3)])

# Calculating phylogenetic signal as well as P-value based on variance of phylogenetically independent contrasts relative to tip shuffling randomization.
physig2 = multiPhylosignal(MatchDataObj$data, MatchDataObj$phy,reps=100000)

physig2

```

## Phylogenetic Signal MCMCglmm models

```{r phylogenetic signal models}

```



# Distribution of Species mean T50 on Phylogeny
```{r phylogenetic tree graph T50}

# Data Table called tip order for Family Tree
TipOrder = data.table(Tip=phy_coll3$tip.label)
# Making a new column that says the order of the tips in the phylogeny
TipOrder[,TipOrder:=.I]

inputData = merge(seedDat_spp,TipOrder,by.x="Species",by.y="Tip",all.x=T)[order(TipOrder)]

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$T50_mean," days)")

p <- ggtree(phy_coll3,layout='rectangular') 
p <- p + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$T50_mean)/10)
p <- p + ggtitle("Species Mean T50")
p
```


# Distribution of Species mean Germination Percent on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$GP,"%)")

p <- ggtree(phy_coll3,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$GP)/10)
p <- p + ggtitle("Species Mean GP")
p
```
# Distribution of Species mean Seed Mapps on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$Seedmass,"mg)")

p <- ggtree(phy_coll3,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=sqrt(inputData$Seedmass/2))
p <- p + ggtitle("Species Mean GP")
p
```
