---
title: "02 Phylogenetic Signal"
author: "Katie Heineman"
date: "6/7/2019"
output: html_document
---

# Analyzing Phylogenetic Signal in Germination Percent and Seed Time to Median Germination

This document also include visualization of the traits on the tree and two versions of Phylogenetic signal:

*Blomberg et al.'s (2003) K*: K is a scaled ratio of the variance among species over the contrasts variance (the latter of which will be low if phylogenetic signal is high)
*Pagel's Lambda*: λ is a scaling parameter for the correlations between species, relative to the correlation expected under Brownian evolution

http://blog.phytools.org/2012/03/phylogenetic-signal-with-k-and.html

```{r setup, include=FALSE}
# Packages
library(data.table)
library(picante)
library(MCMCglmm)

# Actual current version of ggtree not supported in this version of R, but this should work for installation
#source("https://bioconductor.org/biocLite.R")
#biocLite("ggtree")
library(ggtree)

# Ana's seed database
seedDat <- fread("Data/Savanna seed database.csv")
setnames(seedDat,c("Vegetation type", "Dispersal mode","Moisture content", "Seed mass (mg)"), c("VegType", "DM","MC","SM" ))
seedDat[,Dormancy2Cat:=ifelse(Dormancy=="ND","ND","D")]

## Read in Trees 
# Scenario 3:  adding genera or species to their families or genera using the same approach implemented in Phylomatic and BLADJ 
phy_coll3 <- read.tree("Data/phy_coll3.tre")
# Scenario 1:  adding genera or species as basal polytomies within their families or genera 
phy_coll1 <- read.tree("Data/phy_coll1.tre")
# Scenario 2: adding genera or species randomly within their families or genera 
phy_coll2 <- read.tree("Data/phy_coll2.tre")

```

## Data Prep for Phylogenetic Analysis
```{r data prep phylo}
# Prep seed dat by making species with underscore
seedDat[,Species:=gsub(" ","_",Species)]

# Calculating Species Means
seedDat_spp <- seedDat[,.(T50_mean=mean(T50,na.rm=T),log_T50_mean=mean(log(T50),na.rm=T),GP = mean(GP,na.rm=T), sqrt_GP = mean(sqrt(GP),na.rm=T),Seedmass = mean(SM,na.rm=T), Moisture = mean(MC,na.rm=T)), by="Species"]
# The number of sucess is the averager germ percentage X 10 seeds per replicate X 10 reps
seedDat_spp[,Successes:=GP]
seedDat_spp[,Failures:=100-Successes]
seedDat_spp[,Successes:=round( Successes)]
seedDat_spp[,Failures:=round( Failures)]

seedDat_spp1 <- copy(seedDat_spp)
# Make this dataset into a matrix
rowlab <- seedDat_spp1$Species
trait_matrix = as.matrix(seedDat_spp1[,Species:=NULL])
rownames(trait_matrix) = rowlab

```




## Calculating Phylogenetic Signal Using Picante Package

The statistic calculated here is the K statistic with a one tailed test for significance relative to random type reshuffling (https://cran.r-project.org/web/packages/picante/vignettes/picante-intro.pdf)

```{r old school phylogenetic signal, message=FALSE, warning=FALSE}

# Ensuring trait matrix order matches the phylogenetic tree order 
trait_matrix <- trait_matrix[phy_coll3$tip.label,]

# Creating a data object that matches the tree and the traits (only including T50 and GP)
MatchDataObj= match.phylo.data(phy_coll3, trait_matrix[,c(1,3)])

# Calculating phylogenetic signal as well as P-value based on variance of phylogenetically independent contrasts relative to tip shuffling randomization.
physig = multiPhylosignal(MatchDataObj$data, MatchDataObj$phy,reps=100000)
physig


# Testing if the "scenario matters" (Scenario 2 is different than 1 or 3)
# Ensuring trait matrix order matches the phylogenetic tree order 
trait_matrix <- trait_matrix[phy_coll2$tip.label,]

# Creating a data object that matches the tree and the traits (only including T50 and GP)
MatchDataObj= match.phylo.data(phy_coll2, trait_matrix[,c(1,3)])

# Calculating phylogenetic signal as well as P-value based on variance of phylogenetically independent contrasts relative to tip shuffling randomization.
physig2 = multiPhylosignal(MatchDataObj$data, MatchDataObj$phy,reps=100000)

physig2

```

## Phylogenetic Signal MCMCglmm models (Pagel's lambda)

```{r phylogenetic signal models}
# We want to investigate a possible relationship between the phenotypes and the cofactor, 
# while controlling for any phylogenetic dependency in the dataset. 
# To achieve our goal, we want to use the phylogenetic mixed model implemented in the 
# MCMCglmm package. Note the phylo column containing the name of the species in our dataset: 
#   it corresponds to the phylogenetic effect we are going to include in our model. 
# In order to do so, we need to calculate the inverse of the  $ \bm{\Sigma}$ 
#   matrix of phylogenetic correlation:
inv.phylo<-inverseA(phy_coll3,scale=TRUE,nodes="ALL",reduced=FALSE)
inv.phylo$Ainv
Ainv = inv.phylo$Ainv

# Initializing the bayesian model by defining priors
# Here is a prior that has been in several sources that is thought of as a weak prior
prior_weak <- list(R = list(V=1, nu=0.002),
              G = list(G1 = list(V=1, nu=0.002)))
# I changed the nu prior form 0.02 to 2 and it seems to help convergence
prior<-list(G=list(G1=list(V=1,nu=2)),R=list(V=1,nu=0.02))
#  A prior Katie has tinkered with to make stronger
prior_stronger<-list(G=list(G1=list(V=1,nu=0.8)),R=list(V=1,nu=0.8))


# Function calculating a version of pagel's lambda for MCMCglmm objects
lambda <- function(modelObject){
 # objPlot <- plot(modelObject)
  objlambdaVec <- lambda <- modelObject$VCV[,'Species']/
    (modelObject$VCV[,'Species']+modelObject$VCV[,'units'])
  objlambdaMean <- mean(lambda)
  objlambdaMode <- posterior.mode(lambda)
  objlambdaInt <- HPDinterval(lambda)
  
  return(list(lambdaVec=objlambdaVec,lambdaMean=objlambdaMean,lambdaMode=objlambdaMode,lambdaInt=objlambdaInt))
}


#### #### #### #### 
#### T50 #####
#### #### #### #### 
## Bayesian model checking for just phylogenetic signal
## Convergence of this model was helped by converting the T50 variable to a log scale
model_T50<-MCMCglmm(log(T50)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
plot(model_T50)
summary(model_T50)
lambda(model_T50)$lambdaInt

## How much phylogenetic signal in species means?
# Bayesian model checking for just phylogenetic signal
# Convergence of this model was helped by converting the T50 variable to a log scale
model_T50_spp<-MCMCglmm(log(T50_mean)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior_stronger,
                    data=seedDat_spp,nitt=300000,burnin=10000,thin=50)
plot(model_T50_spp)
summary(model_T50_spp)
lambda(model_T50_spp)$lambdaInt

#### #### #### #### 
#### Germiation Percent #####
#### #### #### #### 
# The Bayesian model - including the dependent variable (Germination Percentage)
# prior2=list(R=list(V=1, nu=5), G=list(G1=list(V=1, nu=2,  
#                                              alpha.mu=0, alpha.V=500)))
 prior_to_beat = list(R = list(V = 1, nu=0.5), G = list(G1 = list(V = 1, nu = .5, alpha.mu=0, alpha.V=200)))
 prior_to_beat2 = list(R = list(V = 1, nu=0.1), G = list(G1 = list(V = 1, nu = .5, alpha.mu=0, alpha.V=500)))
 prior1 = list(R = list(V = 1, nu=0.1), G = list(G1 = list(V = 1, nu = 0.8, alpha.mu=0, alpha.V=500)))
 

model_GP_phyloonly<-MCMCglmm(cbind((Successes),(Failures))~1,random=~Species,family="multinomial2",ginverse=list(Species=Ainv),prior = prior1, data=seedDat_spp,nitt=3000000,burnin=100000,thin=500)       
plot(model_GP_phyloonly)
summary(model_GP_phyloonly)
lambda(model_GP_phyloonly)$lambdaInt



prior<-list(G=list(G1=list(V=1,nu=.2)),R=list(V=1,nu=0.02))

model_GP_phyloonly_gauss<-MCMCglmm(GP~1,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior = prior, data=seedDat_spp,nitt=300000,burnin=10000,thin=50)       
plot(model_GP_phyloonly_gauss)
summary(model_GP_phyloonly_gauss)
lambda(model_GP_phyloonly_gauss)$lambdaInt


```



# Distribution of Species mean T50 on Phylogeny
```{r phylogenetic tree graph T50}

# Data Table called tip order for Family Tree
TipOrder = data.table(Tip=phy_coll3$tip.label)
# Making a new column that says the order of the tips in the phylogeny
TipOrder[,TipOrder:=.I]

inputData = merge(seedDat_spp,TipOrder,by.x="Species",by.y="Tip",all.x=T)[order(TipOrder)]

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$T50_mean," days)")

p <- ggtree(phy_coll3,layout='rectangular') 
p <- p + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$T50_mean)/10)
p <- p + ggtitle("Species Mean T50")
p
```


# Distribution of Species mean Germination Percent on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$GP,"%)")

p <- ggtree(phy_coll3,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$GP)/10)
p <- p + ggtitle("Species Mean GP")
p
```
# Distribution of Species mean Seed Mapps on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll3$tip.label <- paste0(inputData$Species, " (" ,inputData$Seedmass,"mg)")

p <- ggtree(phy_coll3,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=sqrt(inputData$Seedmass/2))
p <- p + ggtitle("Species Mean GP")
p
```
