---
title: "02 Bayesian Models with Phylogenetic Random Effects"
author: "Katie Heineman"
date: "6/7/2019"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(lme4)
library(MCMCglmm)

# Ana's seed database
seedDat <- fread("Data/Savanna seed database.csv")
setnames(seedDat,c("Vegetation type", "Dispersal mode","Moisture content", "Seed mass (mg)","Seed life-span"), c("VegType", "DM","MC","SM","LifeSpan"))
seedDat[,Dormancy2Cat:=ifelse(Dormancy=="ND","ND","D")]

# Prep seed dat by making species with underscore
seedDat[,Species:=gsub(" ","_",Species)]

# Create variables ot use for the binomial models
seedDat[,Successes:=GP]
seedDat[,Failures:=100-GP]


## Read in Trees 
# Scenario 3:  adding genera or species to their families or genera using the same approach implemented in Phylomatic and BLADJ 
phy_coll <- read.tree("Data/phy_coll3.tre")
```
## MCMCGlmm Model Set Up

Interesting source that may provide some insight: https://github.com/TGuillerme/mulTree/blob/master/doc/Vanilla_flavoured_phylogenetic_analyses.Rmd

Note: I was confused why the inverse matrix doesn't have each pairwise combination of phylogenetic distances for all species  represented. It seems that Hadfield (2010) argues this is too computationally intensive: https://onlinelibrary.wiley.com/doi/full/10.1111/j.1420-9101.2009.01915.x However, I think maybe I could coerce this through nodes="ALL" argument in inverseA. It seems like these things should not be happening by default but I don't think they are.

Looking for the best model at each covariate level

```{r mcmcglmm model set up, message=FALSE, warning=FALSE}

# We want to investigate a possible relationship between the phenotypes and the cofactor, 
# while controlling for any phylogenetic dependency in the dataset. 
# To achieve our goal, we want to use the phylogenetic mixed model implemented in the 
# MCMCglmm package. Note the phylo column containing the name of the species in our dataset: 
#   it corresponds to the phylogenetic effect we are going to include in our model. 
# In order to do so, we need to calculate the inverse of the  $ \bm{\Sigma}$ 
#   matrix of phylogenetic correlation:
inv.phylo<-inverseA(phy_coll,scale=TRUE,nodes="TIPS",reduced=FALSE)
inv.phylo$Ainv
Ainv = inv.phylo$Ainv
# Evaluating matrix as dataframe
b = as.data.table(summary(Ainv))[order(i,j)]
# Evaluating sample matrix
#Ainv1<-inverseA(bird.families)
#b1 = as.data.table(summary(Ainv1$Ainv))[order(i,j)]
# 


# Function calculating pagel's lambda for MCMCglmm objects
lambda <- function(modelObject){
 # objPlot <- plot(modelObject)
  objlambdaVec <- lambda <- modelObject$VCV[,'Species']/
    (modelObject$VCV[,'Species']+modelObject$VCV[,'units'])
  objlambdaMean <- mean(lambda)
  objlambdaMode <- posterior.mode(lambda)
  objlambdaInt <- HPDinterval(lambda)
  
  return(list(lambdaVec=objlambdaVec,lambdaMean=objlambdaMean,lambdaMode=objlambdaMode,lambdaInt=objlambdaInt))
}

```

## Model selection T50 

```{r model comparison}
## Bayesian model checking for just phylogenetic signal
## Convergence of this model was helped by converting the T50 variable to a log scale
# Initializing the bayesian model by defining priors
# Here is a prior that has been in several sources that is thought of as a weak prior
prior_weak <- list(R = list(V=1, nu=0.002),
              G = list(G1 = list(V=1, nu=0.002)))
# I changed the nu prior form 0.02 to 2 and it seems to help convergence
prior<-list(G=list(G1=list(V=1,nu=0.02)),R=list(V=1,nu=0.02))


## FULL MODEL
model_T50_fullmodel<-MCMCglmm(log(T50)~VegType + DM + MC + log(SM)+Dormancy2Cat+LifeSpan, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=3000000,burnin=100000,thin=500)
plot(model_T50_fullmodel)
summary(model_T50_fullmodel)
lambda(model_T50_fullmodel)$lambdaInt
summary(model_T50_fullmodel)$DIC

## 5 variables (6 chose 5)
# DM + MC + SM + Dorm +  LS
model_T50_5a<-MCMCglmm(log(T50)~ DM + MC + log(SM)+Dormancy2Cat+LifeSpan, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
plot(model_T50_5a)
summary(model_T50_5a)
lambda(model_T50_5a)$lambdaInt
summary(model_T50_5a)$DIC

# DM + MC + SM + Dorm+ VT
model_T50_5b<-MCMCglmm(log(T50)~ DM + MC + log(SM)+Dormancy2Cat+VegType, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

# MC + SM +Dorm +  VT + LS
model_T50_5c<-MCMCglmm(log(T50)~  MC + log(SM)+Dormancy2Cat+VegType + LifeSpan, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

# SM + Dorm + VT + LS + DM  ## BEST MODEL ##
model_T50_5d <-MCMCglmm(log(T50)~  log(SM)+Dormancy2Cat+VegType + LifeSpan + DM, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

# Dorm + VT + LS + DM + MC
model_T50_5e <-MCMCglmm(log(T50)~ Dormancy2Cat+VegType + LifeSpan + DM + MC, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

## 4 variables (5 chose 4 = 4) (Excluding MC variable because not included five factor test)
# DM + LS + SM - Dorm  ## BEST MODEL ##
model_T50_4a <- MCMCglmm(log(T50)~DM + LifeSpan + log(SM)+Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# LS + SM + Dorm + VegType
model_T50_4b <- MCMCglmm(log(T50)~LifeSpan + log(SM)+Dormancy2Cat + VegType ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM + Dorm + VegType + DM
model_T50_4c <- MCMCglmm(log(T50)~ log(SM)+Dormancy2Cat+ VegType + DM ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dorm + VegType + DM + LS 
model_T50_4d <- MCMCglmm(log(T50)~VegType+Dormancy2Cat+DM + LifeSpan ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)

## 3 Variables (4 chose 3 = 4) # Exclude habitat type
# DM + LS + SM 
model_T50_3a <- MCMCglmm(log(T50)~DM + LifeSpan + log(SM) ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
#  LS + SM  + Dorm
model_T50_3b <- MCMCglmm(log(T50)~ LifeSpan + log(SM) + Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM  + Dorm + DM
model_T50_3c <- MCMCglmm(log(T50)~log(SM) + Dormancy2Cat + DM ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dorm + DM + LS
model_T50_3d <- MCMCglmm(log(T50)~ Dormancy2Cat + DM+LifeSpan ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)

## 2 Variables (4 chose 2 = 6) # Keeping both Seed Mass and Dispersal Mode because it was close
# DM + Dorm 
model_T50_2a <- MCMCglmm(log(T50)~DM + Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + SM
model_T50_2b <- MCMCglmm(log(T50)~log(SM)+DM ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM + DORM 
model_T50_2c <- MCMCglmm(log(T50)~log(SM)+Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM + LS
model_T50_2e <- MCMCglmm(log(T50)~log(SM)+LifeSpan ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DORM + LS ## BEST MODEL##
model_T50_2f <- MCMCglmm(log(T50)~LifeSpan+Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + LS
model_T50_2g <- MCMCglmm(log(T50)~LifeSpan+DM ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)

# Interaction of best two
model_T50_2fInt <- MCMCglmm(log(T50)~LifeSpan*Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,data=seedDat,nitt=300000,burnin=10000,thin=50)


## 1 Variable (6)
# DM
model_T50_1a <- MCMCglmm(log(T50)~DM  ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)
# MC
model_T50_1b <- MCMCglmm(log(T50)~MC  ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM
model_T50_1c <- MCMCglmm(log(T50)~ log(SM) ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dorm
model_T50_1d <- MCMCglmm(log(T50)~Dormancy2Cat ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)
# LifeSpan
model_T50_1e <- MCMCglmm(log(T50)~LifeSpan ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)
# VegType
model_T50_1f <- MCMCglmm(log(T50)~VegType ,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior, data=seedDat,nitt=300000,burnin=10000,thin=50)


## Intercept Only 
model_T50<-MCMCglmm(log(T50)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=3000000,burnin=100000,thin=500)
plot(model_T50)
summary(model_T50)
lambda(model_T50)$lambdaInt
summary(model_T50)$DIC


```


# MCMC Glmm Germination Percentage
```{r MCMCglmm GP}
##################################
# Priors (the default priors seem to work almost better)
 prior1 = list(R = list(V = 1, nu=0.002), G = list(G1 = list(V = 1, nu = 0.002, alpha.mu=0, alpha.V=50)))


## FULL MODEL
model_GP_full<-MCMCglmm(cbind(Successes,Failures)~VegType + DM + MC + log(SM)+Dormancy2Cat+ LifeSpan,random=~Species,prior=prior1,
                             family="multinomial2",ginverse=list(Species=Ainv),
                             data=seedDat,nitt=300000,burnin=10000,thin=50)                                         

summary(model_GP_full)
plot(model_GP_full)

## 5 variables
# SM + VegType + Dorm + DM + LS ## BEST MODEL##
model_GP_5 <- MCMCglmm(cbind(Successes,Failures)~log(SM)+VegType+Dormancy2Cat+DM+LifeSpan , random=~Species,family="multinomial2",ginverse=list(Species=Ainv), prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_4)


## 4 variables (5 chose 4 = 4)
# DM + MC + SM + Dorm
model_GP_4a <- MCMCglmm(cbind(Successes,Failures)~DM + MC + log(SM)+Dormancy2Cat , random=~Species,family="multinomial2",ginverse=list(Species=Ainv), prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + MC + SM + VegType
model_GP_4b <- MCMCglmm(cbind(Successes,Failures)~DM + MC + log(SM)+VegType , random=~Species,family="multinomial2",ginverse=list(Species=Ainv), prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# MC + SM + VegType + Dorm
model_GP_4c <- MCMCglmm(cbind(Successes,Failures)~ MC + log(SM)+Dormancy2Cat+ VegType , random=~Species,family="multinomial2",ginverse=list(Species=Ainv), prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM + VegType + Dorm + DM +  ## BEST MODEL##
model_GP_4d <- MCMCglmm(cbind(Successes,Failures)~log(SM)+VegType+Dormancy2Cat+DM , random=~Species,family="multinomial2",ginverse=list(Species=Ainv), prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# VegType + DM + Dorm + LS
model_GP_4e <- MCMCglmm(cbind(Successes,Failures)~VegType+DM+Dormancy2Cat+LifeSpan,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)


## 3 variables (4 chose 3 = 4)
# SM + Vegtype + Dorm
model_GP_3a <- MCMCglmm(cbind(Successes,Failures)~log(SM)+VegType+Dormancy2Cat,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM + VegType + DM 
model_GP_3b <- MCMCglmm(cbind(Successes,Failures)~log(SM)+VegType+DM,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# VegType + DM + Dorm ## BEST MODEL BARELY##s
model_GP_3c <- MCMCglmm(cbind(Successes,Failures)~VegType+DM+Dormancy2Cat,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + Dorm + SM 
model_GP_3d <- MCMCglmm(cbind(Successes,Failures)~DM+Dormancy2Cat+log(SM),random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + Dorm + MC 
model_GP_3e <- MCMCglmm(cbind(Successes,Failures)~DM+Dormancy2Cat+MC,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# DM + Dorm + LS
model_GP_3f <- MCMCglmm(cbind(Successes,Failures)~DM+Dormancy2Cat+LifeSpan,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)


## 2 variables (3 chose 2 = 3)
# VegType + DM ## BEST MODEL##
model_GP_2a <- MCMCglmm(cbind(Successes,Failures)~VegType+DM,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)

#  VegType + Dorm 
model_GP_2b <- MCMCglmm(cbind(Successes,Failures)~VegType+Dormancy2Cat,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dormancy2Cat + DM
model_GP_2c <- MCMCglmm(cbind(Successes,Failures)~Dormancy2Cat+DM,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dormancy2Cat + LifeSpan
model_GP_2d <- MCMCglmm(cbind(Successes,Failures)~Dormancy2Cat+LifeSpan,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)


##1 variables (5)
# DM 
model_GP_1a <- MCMCglmm(cbind(Successes,Failures)~DM,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# SM 
model_GP_1b <- MCMCglmm(cbind(Successes,Failures)~log(SM),random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# VegType 
model_GP_1c <- MCMCglmm(cbind(Successes,Failures)~VegType,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# Dorm ## Best Model Barely ##
model_GP_1d <- MCMCglmm(cbind(Successes,Failures)~Dormancy2Cat,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# MC
model_GP_1e <- MCMCglmm(cbind(Successes,Failures)~MC,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)
# LifeSpan
model_GP_1f <- MCMCglmm(cbind(Successes,Failures)~LifeSpan,random=~Species,family="multinomial2", ginverse=list(Species=Ainv),prior=prior1,data=seedDat,nitt=300000,burnin=10000,thin=50)


## INTERCEPT ONLY
model_GP_phyloonly<-MCMCglmm(cbind((Successes),(Failures))~1,random=~Species,family="multinomial2",ginverse=list(Species=Ainv),prior = prior1, data=seedDat,nitt=300000,burnin=10000,thin=50)                                          


#### Playing with gaussian form for GP
## FULL
model_GP_guass<-MCMCglmm(GP~log(SM)+VegType+Dormancy2Cat+DM+MC,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass)$DIC

## 4 variables 
model_GP_guass4<-MCMCglmm(GP~log(SM)+Dormancy2Cat+DM+MC,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass4)$DIC

## 3 variables 
model_GP_guass3<-MCMCglmm(GP~VegType+Dormancy2Cat+DM,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass3)$DIC


model_GP_guass3a<-MCMCglmm(GP~MC+Dormancy2Cat+DM,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass3a)$DIC


## 2 variables 
model_GP_guass2<-MCMCglmm(GP~Dormancy2Cat+VegType,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass2)$DIC

## 1 variables 
model_GP_guass1<-MCMCglmm(GP~Dormancy2Cat,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)
summary(model_GP_guass1)$DIC

## Looking at intercept only for gaussian model
model_GP_guass_Intonly<-MCMCglmm(GP~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=3000000,burnin=100000,thin=500)
summary(model_GP_guass_Intonly)$DIC
```


## Code for making chains multi-core (may want to explore)

```{r multicore code}

# ## FULL Bayesian model multiple regression model for T50
# setCores<-round(detectCores()*0.8)
# cl <- makeCluster(getOption("cl.cores",setCores))
# cl.pkg <- clusterEvalQ(cl,library(MCMCglmm)) 
# clusterExport(cl,"prior")
#   clusterExport(cl,"test.data")
#   clusterExport(cl,"Ainv")
#  model2_10runs<-parLapply(cl=cl,1:10, function(i) {
#     MCMCglmm(phenotype~1, random=~taxon, ginverse=list(taxon=Ainv),
#              data=test.data, prior=prior, verbose=FALSE, 
#              nitt=100000, burnin=10000, thin=10)}
#   )
#   
# # once it's finished, use stopCluster() to stop running the parallel cluster
# stopCluster(cl)
# summary(model2_10runs[[3]]) # summarize the third model out of the 10
# 
# 
# cl <- makeCluster(getOption("cl.cores",setCores))
# 


```

