---
title: "02 Bayesian Models with Phylogenetic Random Effects"
author: "Katie Heineman"
date: "6/7/2019"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(lme4)
library(MCMCglmm)

# Ana's seed database
seedDat <- fread("Data/Savanna seed database.csv")
setnames(seedDat,c("Vegetation type", "Dispersal mode","Moisture content", "Seed mass (mg)"), c("VegType", "DM","MC","SM" ))
seedDat[,Dormancy2Cat:=ifelse(Dormancy=="ND","ND","D")]

# Prep seed dat by making species with underscore
seedDat[,Species:=gsub(" ","_",Species)]

# Calculating Species Means
seedDat_spp <- seedDat[,.(T50_mean=mean(T50,na.rm=T),log_T50_mean=mean(log(T50),na.rm=T),GP = sum(GP,na.rm=T), sqrt_GP = mean(sqrt(GP),na.rm=T),log_Seedmass = mean(log(SM),na.rm=T), Moisture = mean(MC,na.rm=T)), by="Species"]
# The number of sucess is the averager germ percentage X 10 seeds per replicate X 10 reps
seedDat_spp[,Successes:=GP]
seedDat_spp[,Failures:=1000-Successes]
seedDat_spp[,Successes:=round( Successes)]
seedDat_spp[,Failures:=round( Failures)]


## Read in Trees 
# Scenario 3:  adding genera or species to their families or genera using the same approach implemented in Phylomatic and BLADJ 
phy_coll <- read.tree("Data/phy_coll3.tre")
```
## MCMCGlmm Models for T50 (mean germination time)

Interesting source that may provide some insight: https://github.com/TGuillerme/mulTree/blob/master/doc/Vanilla_flavoured_phylogenetic_analyses.Rmd

Note: I was confused why the inverse matrix doesn't have each pairwise combination of phylogenetic distances for all species  represented. It seems that Hadfield (2010) argues this is too computationally intensive: https://onlinelibrary.wiley.com/doi/full/10.1111/j.1420-9101.2009.01915.x However, I think maybe I could coerce this through nodes="ALL" argument in inverseA. It seems like these things should not be happening by default but I don't think they are.

```{r mcmcglmm T50, message=FALSE, warning=FALSE}

# We want to investigate a possible relationship between the phenotypes and the cofactor, 
# while controlling for any phylogenetic dependency in the dataset. 
# To achieve our goal, we want to use the phylogenetic mixed model implemented in the 
# MCMCglmm package. Note the phylo column containing the name of the species in our dataset: 
#   it corresponds to the phylogenetic effect we are going to include in our model. 
# In order to do so, we need to calculate the inverse of the  $ \bm{\Sigma}$ 
#   matrix of phylogenetic correlation:
inv.phylo<-inverseA(phy_coll,scale=TRUE,nodes="TIPS",reduced=FALSE)
inv.phylo$Ainv
Ainv = inv.phylo$Ainv
# Evaluating matrix as dataframe
b = as.data.table(summary(Ainv))[order(i,j)]
# Evaluating sample matrix
#Ainv1<-inverseA(bird.families)
#b1 = as.data.table(summary(Ainv1$Ainv))[order(i,j)]
# 

# Initializing the bayesian model by defining priors
# Here is a prior that has been in several sources that is thought of as a weak prior
prior_weak <- list(R = list(V=1, nu=0.002),
              G = list(G1 = list(V=1, nu=0.002)))
# I changed the nu prior form 0.02 to 2 and it seems to help convergence
prior<-list(G=list(G1=list(V=1,nu=0.02)),R=list(V=1,nu=0.02))
#  A prior Katie has tinkered with to make stronger
prior_stronger<-list(G=list(G1=list(V=1,nu=0.8)),R=list(V=1,nu=0.8))


# Function calculating pagel's lambda for MCMCglmm objects
lambda <- function(modelObject){
 # objPlot <- plot(modelObject)
  objlambdaVec <- lambda <- modelObject$VCV[,'Species']/
    (modelObject$VCV[,'Species']+modelObject$VCV[,'units'])
  objlambdaMean <- mean(lambda)
  objlambdaMode <- posterior.mode(lambda)
  objlambdaInt <- HPDinterval(lambda)
  
  return(list(lambdaVec=objlambdaVec,lambdaMean=objlambdaMean,lambdaMode=objlambdaMode,lambdaInt=objlambdaInt))
}


## Bayesian model checking for just phylogenetic signal
## Convergence of this model was helped by converting the T50 variable to a log scale
model_T50<-MCMCglmm(log(T50)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=3000000,burnin=100000,thin=500)
plot(model_T50)
summary(model_T50)
lambda(model_T50)$lambdaInt
summary(model_T50)$DIC

# FULL MODEL
model_T50_fullmodel<-MCMCglmm(log(T50)~VegType + DM + MC + log(SM)+Dormancy2Cat, random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=3000000,burnin=100000,thin=500)
plot(model_T50_fullmodel)
summary(model_T50_fullmodel)
lambda(model_T50_fullmodel)$lambdaInt
summary(model_T50_fullmodel)$DIC

model_T50_dormancyCatOnly <- MCMCglmm(log(T50)~Dormancy2Cat,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

plot(model_T50_dormancyCatOnly)
summary(model_T50_dormancyCatOnly)
lambda(model_T50_dormancyCatOnly)$lambdaInt
summary(model_T50_dormancyCatOnly)$DIC

```


# MCMC Glmm Germination Percentage
```{r MCMCglmm GP}
##################################
########
# The Bayesian model - including the dependent variable (Germination Percentage)
 prior2=list(R=list(V=1, nu=50), G=list(G1=list(V=1, nu=2,  
                                               alpha.mu=0, alpha.V=500)))
 prior1 = list(R = list(V = 1, nu=0.08), G = list(G1 = list(V = 1, nu = .5, alpha.mu=0, alpha.V=200)))

seedDat_spp[,Successes:=round( Successes)]
 seedDat_spp[,Failures:=round( Failures)]

model_GP_phyloonly<-MCMCglmm(cbind((Successes),(Failures))~1,random=~Species,family="multinomial2",ginverse=list(Species=Ainv),prior = prior1, data=seedDat_spp,nitt=2600000,burnin=60000,thin=200)                                          

plot(model_GP_phyloonly)
summary(model_GP_phyloonly)
lambda_GP <- model_GP_phyloonly $VCV[,'Species']/
  (model_GP_phyloonly$VCV[,'Species']+ model_GP_phyloonly$VCV[,'units'])
mean(lambda_GP)
posterior.mode(lambda_GP)
HPDinterval(lambda_GP)
DICGP <- summary(model_GP_phyloonly)$DIC
#

model_GP_full<-MCMCglmm(cbind(Success,Failure)~DM+habitat+Lifeform+Lifespan,random=~Species,
                             family="multinomial2",ginverse=list(Species=Ainv),prior=prior2,
                             data=data_lim_spp,nitt=5000000,burnin=100000,thin=500)                                          
plot(model_GP_full)
summary(model_GP_full)
lambda_GPfull <- model_GP_full $VCV[,'Species']/
  (model_GP_full$VCV[,'Species']+ model_GP_full$VCV[,'units'])
mean(lambda_GPfull)
posterior.mode(lambda_GPfull)
HPDinterval(lambda_GPfull)
DICfull <- summary(model_GP_full)$DIC

```


## Code for making chains multi-core (may want to explore)

```{r multicore code}

# ## FULL Bayesian model multiple regression model for T50
# setCores<-round(detectCores()*0.8)
# cl <- makeCluster(getOption("cl.cores",setCores))
# cl.pkg <- clusterEvalQ(cl,library(MCMCglmm)) 
# clusterExport(cl,"prior")
#   clusterExport(cl,"test.data")
#   clusterExport(cl,"Ainv")
#  model2_10runs<-parLapply(cl=cl,1:10, function(i) {
#     MCMCglmm(phenotype~1, random=~taxon, ginverse=list(taxon=Ainv),
#              data=test.data, prior=prior, verbose=FALSE, 
#              nitt=100000, burnin=10000, thin=10)}
#   )
#   
# # once it's finished, use stopCluster() to stop running the parallel cluster
# stopCluster(cl)
# summary(model2_10runs[[3]]) # summarize the third model out of the 10
# 
# 
# cl <- makeCluster(getOption("cl.cores",setCores))
# 


```

