---
title: "Savannah Germination"
author: "Katie Heineman"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(lme4)
library(car)
library(ggplot2)
library(ape)
library(MCMCglmm)
library(phytools)
library(phylolm)
library(stringr)
library(readr)

## Making the tree- Doesn't seem to work with Ubuntu this version because it doesn't have the right version of R
# if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
# BiocManager::install()
#biocLite("ggtree")
#library(ggtree)


# Ana's seed database
seedDat <- fread("Data/Savanna seed database.csv")
setnames(seedDat,c("Vegetation type", "Dispersal mode","Moisture content", "Seed mass (mg)"), c("VegType", "DM","MC","SM" ))
seedDat[,Dormancy2Cat:=ifelse(Dormancy=="ND","ND","D")]

```


# Building New Phylogeny using PhyloMakeR

S.Phylomater produces phylogenies using three scenarios. This function & method were described in Qian & Jin 2016 in Journal of Ecology.

1. adding genera or species as basal polytomies within their families or genera 
2. adding genera or species randomly within their families or genera 
3. adding genera or species to their families or genera using the same approach implemented in Phylomatic and BLADJ 

There is also a newer function called VPhyloMaker is supposed to be faster and perhaps use a more up to date phylogeny?

```{r new phylogeny}
## Practice using the tool with example data
# Source the code from Qian & Jin 2016 paper
source("R_codes for S.PhyloMaker.R")
# Read in the example list
# example_splist <- na.omit(as.data.frame(read_delim("example.splist.csv", "\t", escape_double = FALSE, trim_ws = TRUE)))
# # Read in the megaphylogeny
 phylo<-read.tree("Data/PhytoPhylo.tre") # read in the megaphylogeny.
 nodes <- as.data.table(read_delim("Data/nodes.csv", "\t", escape_double = FALSE,  trim_ws = TRUE))
# # Run the phylogeny for two "scenarios"
# result<-S.PhyloMaker(spList=example_splist, tree=phylo, nodes=nodes, scenarios=c("S1","S2")) # For some reason scenario S3 doesn't work

## Making species list for Ana's species
Spp <- seedDat[,.(Count=.N),by="Species"]
Spp[,genus:=word(Species,1,1),by="Species"]

## Finding Family's from genera from phylomaker Nodes file
genusTable = nodes[,.(family=family[1]),by="genus"]

## Merge families onto genus
Spp <- merge(Spp,genusTable,by="genus",all.x=T)
Spp[genus=="Eremanthus", family:="Asteraceae"]
setnames(Spp,"Species","species")
Spp <- Spp[,.(species,genus,family)]

## Bulding Tree for all three scenarios
result_ana_S3 <- S.PhyloMaker(spList=as.data.frame(Spp), tree=phylo, nodes=nodes, scenarios=c("S1","S2","S3")) 

## Comparing scenarios
plot(result_ana_S3$Scenario.1, main="Scenario 1 (basal polytomies)")
plot(result_ana_S3$Scenario.2,main="Scenario 2 (random placement)")
plot(result_ana_S3$Scenario.3,main="Scenario 3 (Bladj & Phylocom)")



```

## Model for T50 (mean germination time)

Interesting source that may provide some insight: https://github.com/TGuillerme/mulTree/blob/master/doc/Vanilla_flavoured_phylogenetic_analyses.Rmd

```{r mcmcglmm T50, message=FALSE, warning=FALSE}
# Prep seed dat by making species with underscore
seedDat[,Species:=gsub(" ","_",Species)]

# Species Means
# Calculate species mean T50
seedDat_spp <- seedDat[,.(T50_mean=mean(T50,na.rm=T), GP = mean(GP,na.rm=T), Seedmass = mean(SM,na.rm=T), Moisture = mean(MC,na.rm=T)), by="Species"]
# The number of sucess is the averager germ percentage X 10 seeds per replicate X 10 reps
seedDat_spp[,Successes:=GP/100*10*10]
seedDat_spp[,Failures:=100-Successes]


# Collapsing singles to make this work better (using scenario 3 - BLADJ/PHylocom)
phy_coll <- collapse.singles(result_ana_S3$Scenario.1)
# Naming the internal nodes so that the following inverse function works
phy_coll$node.label[which(phy_coll$node.label=="")]=paste0("unknownNode",seq(1:(length(which(phy_coll$node.label=="")))))

# This does somethign to make the tree ultrametric
phy_coll <- chronoMPL(phy_coll)

# We want to investigate a possible relationship between the phenotypes and the cofactor, 
# while controlling for any phylogenetic dependency in the dataset. 
# To achieve our goal, we want to use the phylogenetic mixed model implemented in the 
# MCMCglmm package. Note the phylo column containing the name of the species in our dataset: 
#   it corresponds to the phylogenetic effect we are going to include in our model. 
# In order to do so, we need to calculate the inverse of the  $ \bm{\Sigma}$ 
#   matrix of phylogenetic correlation:
inv.phylo<-inverseA(phy_coll,nodes="TIPS",scale=TRUE)
inv.phylo$Ainv
Ainv = inv.phylo$Ainv

# Initializing the bayesian model by defining priors
# Here is a prior that has been in several sources that is thought of as a weak prior
prior_weak <- list(R = list(V=1, nu=0.002),
              G = list(G1 = list(V=1, nu=0.002)))
# I changed the nu prior form 0.02 to 2 and it seems to help convergence
prior<-list(G=list(G1=list(V=1,nu=.02)),R=list(V=1,nu=0.02))


## Bayesian model checking for just phylogenetic signal
## Convergence of this model was helped by converting the T50 variable to a log scale
model_T50<-MCMCglmm(log(T50)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=300000,burnin=10000,thin=50)

plot(model_T50)
summary(model_T50)
lambda <- model_T50$VCV[,'Species']/
  (model_T50$VCV[,'Species']+model_T50$VCV[,'units'])
mean(lambda)
posterior.mode(lambda)
HPDinterval(lambda)
DICint <- summary(model_T50)$DIC


#  A prior Katie has tinkered with to make stronger
prior_stronger<-list(G=list(G1=list(V=1,nu=.5)),R=list(V=1,nu=0.1))


## How much phylogenetic signal in species means?
# Bayesian model checking for just phylogenetic signal
# Convergence of this model was helped by converting the T50 variable to a log scale
model_T50_spp<-MCMCglmm((T50_mean)~1,random=~Species,
                    family="gaussian",ginverse=list(Species=Ainv),prior=prior_stronger,
                    data=seedDat_spp,nitt=3000000,burnin=10000,thin=50)


plot(model_T50_spp)
summary(model_T50_spp)
lambda <- model_T50_spp$VCV[,'Species']/
  (model_T50_spp$VCV[,'Species']+model_T50_spp$VCV[,'units'])
mean(lambda)
posterior.mode(lambda)
HPDinterval(lambda)
DICint <- summary(model_T50_spp)$DIC


## FULL Bayesian model multiple regression model for T50
setCores<-round(detectCores()*0.8)
cl <- makeCluster(getOption("cl.cores",setCores))
cl.pkg <- clusterEvalQ(cl,library(MCMCglmm)) 
clusterExport(cl,"prior")
  clusterExport(cl,"test.data")
  clusterExport(cl,"Ainv")
 model2_10runs<-parLapply(cl=cl,1:10, function(i) {
    MCMCglmm(phenotype~1, random=~taxon, ginverse=list(taxon=Ainv),
             data=test.data, prior=prior, verbose=FALSE, 
             nitt=100000, burnin=10000, thin=10)}
  )
  
# once it's finished, use stopCluster() to stop running the parallel cluster
stopCluster(cl)
summary(model2_10runs[[3]]) # summarize the third model out of the 10


cl <- makeCluster(getOption("cl.cores",setCores))



model_T50_fullmodel<-MCMCglmm(log(T50)~VegType + DM + MC + SM+Dormancy2Cat,random=~Species,family="gaussian",ginverse=list(Species=Ainv),prior=prior,
                    data=seedDat,nitt=5000000,burnin=100000,thin=500)
plot(model_T50_fullmodel)
summary(model_T50_fullmodel)
lambda_full <- model_T50_fullmodel$VCV[,'Species']/
  (model_T50_fullmodel$VCV[,'Species']+model_T50_fullmodel$VCV[,'units'])
mean(lambda_full)
posterior.mode(lambda_full)
HPDinterval(lambda_full)
DICfull <- summary(model_T50_fullmodel)$DIC


```
# MCMC Glmm Germination Percentage
```{r MCMCglmm GP}
##################################
########
# The Bayesian model - including the dependent variable (Germination Percentage)
 prior2=list(R=list(V=1, nu=50), G=list(G1=list(V=1, nu=2,  
                                               alpha.mu=0, alpha.V=500)))
 prior1 = list(R = list(V = 1, nu=0.08), G = list(G1 = list(V = 1, nu = .5, alpha.mu=0, alpha.V=200)))

seedDat_spp[,Successes:=round( Successes)]
 seedDat_spp[,Failures:=round( Failures)]

model_GP_phyloonly<-MCMCglmm(cbind((Successes),(Failures))~1,random=~Species,family="multinomial2",ginverse=list(Species=Ainv),prior = prior1, data=seedDat_spp,nitt=2600000,burnin=60000,thin=200)                                          

plot(model_GP_phyloonly)
summary(model_GP_phyloonly)
lambda_GP <- model_GP_phyloonly $VCV[,'Species']/
  (model_GP_phyloonly$VCV[,'Species']+ model_GP_phyloonly$VCV[,'units'])
mean(lambda_GP)
posterior.mode(lambda_GP)
HPDinterval(lambda_GP)
DICGP <- summary(model_GP_phyloonly)$DIC
#

model_GP_full<-MCMCglmm(cbind(Success,Failure)~DM+habitat+Lifeform+Lifespan,random=~Species,
                             family="multinomial2",ginverse=list(Species=Ainv),prior=prior2,
                             data=data_lim_spp,nitt=5000000,burnin=100000,thin=500)                                          
plot(model_GP_full)
summary(model_GP_full)
lambda_GPfull <- model_GP_full $VCV[,'Species']/
  (model_GP_full$VCV[,'Species']+ model_GP_full$VCV[,'units'])
mean(lambda_GPfull)
posterior.mode(lambda_GPfull)
HPDinterval(lambda_GPfull)
DICfull <- summary(model_GP_full)$DIC

```

# Distribution of Species mean T50 on Phylogeny
```{r phylogenetic tree graph T50}

# Data Table called tip order for Family Tree
TipOrder = data.table(Tip=phy_coll$tip.label)
# Making a new column that says the order of the tips in the phylogeny
TipOrder[,TipOrder:=.I]

inputData = merge(seedDat_spp,TipOrder,by.x="Species",by.y="Tip",all.x=T)[order(TipOrder)]

phy_coll$tip.label <- paste0(inputData$Species, " (" ,inputData$T50_mean," days)")

p <- ggtree(phy_coll,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$T50_mean)/10)
p <- p + ggtitle("Species Mean T50")
p
```

# Distribution of Species mean Germination Percent on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll$tip.label <- paste0(inputData$Species, " (" ,inputData$GP,"%)")

p <- ggtree(phy_coll,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=(inputData$GP)/10)
p <- p + ggtitle("Species Mean GP")
p
```
# Distribution of Species mean Seed Mapps on Phylogeny
```{r phylogenetic tree graph T50}

phy_coll$tip.label <- paste0(inputData$Species, " (" ,inputData$Seedmass,"mg)")

p <- ggtree(phy_coll,layout='rectangular') + geom_tiplab( size=3, color="black",offset=3)
p <- p + xlim(0,200)
p <- p +geom_tippoint(aes(x=x), size=sqrt(inputData$Seedmass/2))
p <- p + ggtitle("Species Mean GP")
p
```
